Instant Continuation Marks--Just Add $\lambda$

### 2012-10-10 paper.tex
Winding it up.

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{slatex}
\usepackage{color}

@@

Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection. Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail} and to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}. [Mention parameters?]

@@

We now address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv.

We can express the \scheme'let' construct in \lv\ with application. An expression such as
\begin{schemedisplay}
(let ((x_1 e_1)
      ...
      (x_n e_n))
  e)
\end{schemedisplay}
can be interpreted as
\begin{schemedisplay}
(...((lambda (x_1) ... (lambda (x_n)
           e)...) e_1) ...) e_n)
\end{schemedisplay}
which is the curried form of
\begin{schemedisplay}
((lambda (x_1 ... x_n)
           e) e_1 ... e_n)
\end{schemedisplay}
We unfold this characterization of \scheme'let' to guide the construction of $\mathcal{C}$ before simplifying.
[Drop everything after ``with application.''?]

To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig. \ref{church-encodings}.

@@

\begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
\begin{schemedisplay}
(lambda (kont)
  (lambda (flag)
    (lambda (marks)
      (kont (lambda (x) Ccps[e])))))
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Ccps[x]|}
\begin{schemedisplay}
(lambda (kont)
  (lambda (flag)
    (lambda (marks)
      (kont x))))
\end{schemedisplay}
\end{schemedefn}

We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.

\begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
\begin{schemedisplay}
(lambda (x) Ccps[e])
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Chcps[p]|}
\begin{schemedisplay}
(((Ccps[p] (lambda (x) x)) FALSE) Cpcps[nil])
\end{schemedisplay}
\end{schemedefn}

@@

Now consider the reduction of the same program transformed. We apply the transformation just-in-time as we reduce to reduce term size and promote clarity.
\begin{schemedisplay}
Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]

((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])

@@

[This example is way too long. Are trivial examples generally uninteresting?]

@@

We have discovered a truly marvelous proof of this, which this paper is too short to contain.

We offer a sketch of the proof here.

@@

%\spnewtheorem*{maintheoremeu}{Correctness of $\hat{\mathcal{C}}$}{a}{x}
\begin{theorem}
For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
\end{theorem}

We do this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate contexts which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We first define
\begin{plaindef}
\[
\xi(E)=\begin{cases}
\mathbf{true} &\text{if $E=E'[\wcm{v'}{\hole}]$ for some $E'$ and $v'$}\\
\mathbf{false} &\text{otherwise}
\end{cases}
\]
\end{plaindef}
to denote the \scheme'flags' argument and assume that the \scheme'marks' argument is \scheme|Cpcps[chiE]|. We can now define $\mathcal{C}_{\mathrm{cps}}$ over contexts:

\begin{schemedefn}{\scheme|Ccps[hole]|}
\begin{schemedisplay}
(lambda (value)
  value)
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Ccps[E[(hole rand-value)]]|}
\begin{schemedisplay}
(lambda (rator-value)
  (((Ccps[rand-expr]
     (lambda (rand-value)
       ((((rator-value rand-value) Ccps[E]) xiE) Cpcps[chiE])))
    FALSE)
  Cpcps[chiE]))
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Ccps[E[(v_0 hole)]]|}
\begin{schemedisplay}
(lambda (rand-value)
  ((((v_0 rand-value) Ccps[E]) xiE) Cpcps[chiE]))
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Ccps[E[(wcm hole body-expr)]]|}
\begin{schemedisplay}
(lambda (mark-value) 
  ((lambda (rest-marks) 
      (((Ccps[body-expr] Ccps[E]) TRUE) Chcps[((CONS mark-value) rest-marks)]))
    ((xiE Chcps[(SND chiE)]) Cpcps[chiE])))
\end{schemedisplay}
\end{schemedefn}

\begin{schemedefn}{\scheme|Ccps[E[(wcm v_0 hole)]]|}
\begin{schemedisplay}
Ccps[E]
\end{schemedisplay}
\end{schemedefn}

This allows us to define $\mathcal{C}_{\mathrm{cps}}$ over a context-expression pair with mildly unorthodox notation.

\begin{schemedefn}{\scheme|Ccps[E[e]]|}
\begin{schemedisplay}
(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
\end{schemedisplay}
\end{schemedefn}

From this definition, it is apparent that \scheme|Chcps[p]|=\scheme|(((Ccps[p] Ccps[hole]) XI(hole)) Cpcps[hole])|=\scheme|Ccps[hole[p]]|.

This in turn allows us to express the following lemmas.

\begin{lemma}
For all contexts $E$ and terms $e$, $E[e]\cmrrs v\implies\C{E[e]}\lvrrs\Cp{v}$.
\end{lemma}

To accomplish this, we first prove by structural induction that for all terms $e$ and context $E$ arbitrary but fixed, $E[e]\cmrrs E[v]\implies\C{E[e]}\lvrrs\C{E[v]}$.

We then prove that for all contexts $E$, $E[v]\cmrrs v'\implies\C{E[v]}\lvrrs\Cp{v'}$.

The theorem follows.

@@

Expressing continuation marks in a monadic way is simple enough. In Haskell,
\begin{verbatim}
data CM m a = CM ((Bool,[m]) -> a)

@@

runCM :: CM m a -> a
runCM (CM m) = m (False,[])
\end{verbatim}
is one way, using our approach. The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad. Specifically, the right identity law
\begin{verbatim}
m >>= return = m
\end{verbatim}
does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}). Ager et al. \cite{ager2003functional} implemented a similar stack-inspection monad in terms of a lifted state monad, but resorted to manual management of the stack frames.

\section{Future Work}

@@

\section{Conclusion}

>>>>

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{slatex}
\usepackage{listings}

@@

Continuation marks \cite{clements2006portable} are a programming language feature which generalizes stack inpection. Not only do they dramatically simplify correct instrumentation \cite{clements2001modeling}, they have been used to allow inspection-based dynamic security checks in the presence of tail-call optimization \cite{clements2004tail}, to express aspect-oriented programming in higher-order languagues \cite{tucker2003pointcuts}, and to implement a form of dynamic binding in parameters.

@@

Now, to address the absence of \scheme'let', \scheme'if', \scheme'cons', etc. from \lv: We can express the \scheme'let' construct in \lv\ with application. To achieve \scheme'if' and conditionals as well as list primitives \scheme'cons', \scheme'snd', and \scheme'nil', we use the Church encodings of fig. \ref{church-encodings}.

@@

\noindent
\begin{minipage}{0.5\linewidth}
\begin{schemedefn}{\scheme|Ccps[v]|=\scheme|Ccps[(lambda (x) e)]|}
\begin{schemedisplay}
(lambda (kont)
  (lambda (flag)
    (lambda (marks)
      (kont (lambda (x) Ccps[e])))))
\end{schemedisplay}
\end{schemedefn}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\begin{schemedefn}{\scheme|Ccps[x]|}
\begin{schemedisplay}
(lambda (kont)
  (lambda (flag)
    (lambda (marks)
      (kont x))))
\end{schemedisplay}
\end{schemedefn}
\end{minipage}
\\
\\

We include corresponding definitions for $\mathcal{C}'$ and $\hat{\mathcal{C}}$.
\\
\\
\noindent
\begin{minipage}{0.5\linewidth}
\begin{schemedefn}{\scheme|Cpcps[(lambda (x) e)]|}
\begin{schemedisplay}
(lambda (x) Ccps[e])
\end{schemedisplay}
\end{schemedefn}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\begin{schemedefn}{\scheme|Chcps[p]|}
\begin{schemedisplay}
(((Ccps[p] (lambda (x) x)) FALSE) Cpcps[nil])
\end{schemedisplay}
\end{schemedefn}
\end{minipage}

@@

Now consider the reduction of the same program transformed. We apply the transformation just-in-time as we reduce to prevent term size explosion and promote clarity and omit the uninteresting reductions.
\begin{schemedisplay}
Ch[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))]
; By definition this is
((C[(wcm 0 ((lambda (x) (wcm x (ccm))) 1))] FALSE) Cp[NIL])

@@

@@

Because of space limitations, we offer only a sketch of the proof here. [The full proof can be found in a technical report at ...]

@@

\begin{theorem}
For all programs $p\in\lambda_{cm}$, $\Chcps{\mathrm{eval}_{cm}(p)}=\mathrm{eval}_{v}(\Chcps{p})$.
\end{theorem}

We prove this by overloading $\mathcal{C}_{\mathrm{cps}}$ to accommodate first evaluation contexts and then context-term pairs (denoted in unorthodox style) by
\begin{schemedefn}{\scheme|Ccps[E[e]]|}
\begin{schemedisplay}
(((Ccps[e] Ccps[E]) xiE) Cpcps[chiE])
\end{schemedisplay}
\end{schemedefn}
\noindent
which allows us to formally relate $E[e]$ and $\Ccps{E[e]}$. We then prove a lemma
\begin{lemma}[Simulation]
For all contexts $E\in\lambda_{cm}$ and expressions $e\in\lambda_{cm}$, $E[e]\cmrr E'[e']\implies\C{E[e]}\lvrrs\C{E'[e']}$
\end{lemma}
by induction over contexts $E$ and terms $e$. This lemma justifies the evolution of the evaluation context and each rule in the reduction relation except rule \ref{reduction}. We express this with
\begin{lemma}[Reduction]
For all values $v_0,v_1\in\lambda_{cm}$, $\app{v_0}{v_1}\cmrr e'\implies\app{\Cpcps{v_0}}{\Cpcps{v_1}}\lvrr\Ccps{e'}$.
\end{lemma}

With these lemmas, the theorem follows quickly.

@@

Expressing continuation marks in a monadic way is simple enough. In Haskell,
\begin{lstlisting}[language=Haskell]
data CM m a = CM ((Bool,[m]) -> a)

@@

runCM :: CM m a -> a
runCM (CM m) = m (False,[])
\end{lstlisting}
is one way, using our approach. The factorial functions leveraging this implementation exhibit the correct continuation mark behavior but this is \emph{not} a monad. Specifically, the right identity law, \texttt{m >>= return = m}, does not hold since \texttt{CM} conceptually adds a stack frame in bind (\texttt{>>=}). Ager et al. \cite{ager2003functional} implemented a similar stack-inspection monad in terms of a lifted state monad, but resorted to manual management of the stack frames.

\section{Future Work and Conclusion}

@@
